Laplacian envelope

### 2019-03-30 Notes/laplace.tex
typo
removed some slides

Given a sequence of plasticity events at times $t_1,t_2,t_3,\ldots$, the probability distribution across the internal states at time $t_n$ is
%
\begin{equation}\label{eq:plastseq}
  \pr(t_n) = \pr(t_0)  \M\pot \M\dep \M\dep \M\pot \M\dep \M\pot \M\pot \M\dep \ldots \M\pot .
\end{equation}
%
If the probability of a given plasticity event being poetitiating/depressing os $f\potdep$, the probability of the sequence above (conditioned on the number of events) is given by $f\pot f\dep f\dep f\pot f\dep f\pot f\pot f\dep \ldots f\pot$.

@@

This results in
%
\begin{equation}\label{eq:overlapwt}
\begin{aligned}
  \av{\syn^1_\text{ideal}\syn^1(t)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \e^{rt\frg} \w, \\
  \av{\syn^1_\text{ideal}\syn^1(\infty)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \onev\eq \w \\
         &= \eq (f\pot\onev-f\dep\onev)\, \eq \w \\
         &=  (f\pot-f\dep)\, \eq \w \\
         &=  (f\pot-f\dep)\, \eq \e^{rt\frg} \w .
\end{aligned}
\end{equation}
%
Combining these allows us to write the numerator of \eqref{eq:SNRdef} as
%
\begin{equation}\label{eq:signal}
\begin{aligned}
  \av{\syn_\text{ideal}\cdt\syn(t)} - \av{\syn_\text{ideal}\cdt\syn(\infty)}
    &= N \eq \enc \, \e^{(r\frg+\hom)t} \w , \\
    \text{where} \qquad
  \enc &= f\pot(\M\pot-\I)-f\dep(\M\dep-\I)
\end{aligned}
\end{equation}
%
Combining with \eqref{eq:noise} gives
%
\begin{equation}\label{eq:SNRcurveExact}
  \SNR(t) = \frac{\sqrt{N} \eq \enc\,\e^{(r\frg+\hom)t}\w}
                 {\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
\end{equation}
%
We can absorb $\hom$ into $\MM\potdep$ by defining
%
\begin{equation}\label{eq:absorbhom}
\begin{aligned}
  \Mh\potdep &= \M\potdep + \frac{\hom}{2rf\potdep},
  &\qquad
  \frgh &= f\pot\Mh\pot + f\dep\Mh\dep - \I,
  \\&&
  \ench &= f\pot(\Mh\pot-\I)-f\dep(\Mh\dep-\I).
\end{aligned}
\end{equation}
%
Then the SNR curve is
%
\begin{equation}\label{eq:SNRcurvehom}
  \SNR(t) = \frac{\sqrt{N}  \eq \ench \,\e^{rt\frgh}\w}
                 {\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
\end{equation}
%
The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with $f\pot=f\dep$ or $\eq_+=\eq_-$.
We can perform the maximisation in two steps.
First maximise the numerator at fixed $\eq\w$.
Then maximise the ratio \wrt $\eq\w$.
This will allow us to ignore the denominator until the very end.

We can then treat the off-diagonal elements of $\Mh\potdep$ as the independent degrees of freedom, subject to these constraints, when maximising the SNR.
The diagonal elements are determined by setting the row sums to 1.
We then have
%
\begin{equation}\label{eq:derivpd}
  \pdiff{\M\pot_{ij}}{\M\pot_{mn}} = \pdiff{\M\dep_{ij}}{\M\dep_{mn}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
  \qquad
  \pdiff{\M\pot_{ij}}{\M\dep_{mn}} = \pdiff{\M\dep_{ij}}{\M\pot_{mn}} = 0.
\end{equation}
%
It will also be convenient to define the differential operators
%
\begin{equation}\label{eq:pertfe}
  \pdiff{}{\frg_{ij}} = \frac{1}{2f\pot} \pdiff{}{\M\pot_{ij}} + \frac{1}{2f\dep} \pdiff{}{\M\dep_{ij}},
  \qquad
  \pdiff{}{\enc_{ij}} = \frac{1}{2f\pot} \pdiff{}{\M\pot_{ij}} - \frac{1}{2f\dep} \pdiff{}{\M\dep_{ij}},
\end{equation}
%
these satisfy
%
\begin{equation}\label{eq:derivfe}
  \pdiff{\frg_{ij}}{\frg_{mn}} = \pdiff{\enc_{ij}}{\enc_{mn}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
  \qquad
  \pdiff{\frg_{ij}}{\enc_{mn}} = \pdiff{\enc_{ij}}{\frg_{mn}} = 0.
\end{equation}
%

@@

Now consider the Laplace transform of the SNR curve \eqref{eq:SNRcurvehom} (ignoring the denominator)
%
\begin{equation}\label{eq:lsnr}
\begin{aligned}
  A(s) &= \intd[_0^{\infty}]{t} \e^{-st}\SNR(t) \\
   &= \intd[_0^{\infty}]{t} \frac{ \sqrt{N} \eq \enc \e^{(r\frg-s)t} \w }{ \sqrt{1-(f\pot-f\dep)^2(\eq\w)^2} } \\
   &= \frac{\sqrt{N} \eq \enc \prn{s - r\frg}\inv \w }{\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
\end{aligned}
\end{equation}
%
This expression is ill-behaved at $s=0$.
Thanks to \eqref{eq:alfund}, we can solve this by the replacement $\mathbf{G}(s) \to \fund(s)$, as $\enc\onev=0$.
%
\begin{equation}\label{eq:larea}
\begin{aligned}
  \hat{A}(s) = A(s) \sqrt{\frac{{{1-(f\pot-f\dep)^2(\eq\w)^2}}}{N}}  &=  \eq \enc \fund(s) \w \\
    &= \sum_{ijk} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \eq_k \w_k \\
    &= \sum_{ij}  \eq_i \enc_{ij} (\etw_{i}(s) - \etw_{j}(s)).
\end{aligned}
\end{equation}
%
Note that $A(0)=A$, the area, and $\lim_{s\to\infty} \brc{sA(s)} = \SNR(0)$, the initial SNR.

In analogy with the case $s=0$, we define
%
\begin{equation}\label{eq:lareacoeffs}
  \begin{aligned}
    c_k(s) &= \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \\
    a_i(s) &= \sum_j \eq_i \enc_{ij} (\etw_{i}(s) - \etw_{j}(s)),\\
    \implies
    \hat{A}(s) &= \sum_k c_k(s) \eq_k \w_k
      = \sum_i a_i(s).
  \end{aligned}
\end{equation}
%

@@

Using the same approach as the area bound in \cite{Lahiri2013synapse}, and defining $\thb_i(s) = \sum_j \fptb_{ij}(0) a_j(s)$,  we find that the derivatives are
%
\begin{equation}\label{eq:lAdiff}
  \pdiff{\hat{A}(s)}{\M\potdep_{mn}}
     = f\potdep \eq_m \prn{ r \brk{\thb_m(s) - \thb_n(s)}
     +\prn{rc_m(s) \pm 1} \brk{\etw_m(s) - \etw_n(s)} },
\end{equation}
%
with the upper/lower sign in $\pm$ corresponding to derivatives \wrt $\M\potdep$.
The term with $\Theta(s)$ comes from differentiating $\eq$, the term with $c(s)$ comes from differentiating $\etw(s)$ and the term with $\pm1$ comes from differentiating $\enc$ in \eqref{eq:larea}.

To write down the Hessian, we will first define some notation.
Given a four-index array $H_{ijkl}$, we define $H\trans_{ijkl} = H_{klij}$
Then we denote
%
\begin{equation}\label{eq:hessparts}
\begin{aligned}
  R_{ijkl} &=
    &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\thb_k(s) - \thb_l(s)} \\
    &&&+ \eq_i \brc{[\fund(0)\enc\fund(s)]_{ik} - [\fund(0)\enc\fund(s)]_{jk}} \brc{\etw_k(s) - \etw_l(s)} \\
    &&&+ \eq_i c_i(s) \brc{\fund_{ik}(s) - \fund_{jk}(s)} \brc{\etw_k(s) - \etw_l(s)},  \\
  S_{ijkl} &=
    &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\etw_k(s) - \etw_l(s)} \\
    &&&+ \eq_k \brc{\fund_{ki}(s) - \fund_{li}(s)} \brc{\etw_i(s) - \etw_j(s)}.
\end{aligned}
\end{equation}
%
Then we can write the hessian as
%
\begin{equation}\label{eq:hessian}
\begin{aligned}
  \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\pot_{kl}} &=
     (f\pot)^2 (R_{ijkl} + R\trans_{ijkl})
     + f\pot(S_{ijkl} + S\trans_{ijkl}), \\
  \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\dep_{kl}} &=
     f\pot f\dep (R_{ijkl} + R\trans_{ijkl})
     - f\pot S_{ijkl} + f\dep S\trans_{ijkl}, \\
  \pdiff{^2\hat{A}(s)}{\M\dep_{ij} \partial\M\pot_{kl}} &=
     f\pot f\dep (R_{ijkl} + R\trans_{ijkl})
     - f\pot S\trans_{ijkl} + f\dep S_{ijkl}, \\
  \pdiff{^2\hat{A}(s)}{\M\dep_{ij} \partial\M\dep_{kl}} &=
     (f\dep)^2 (R_{ijkl} + R\trans_{ijkl})
     - f\dep(S_{ijkl} + S\trans_{ijkl}), \\
\end{aligned}
\end{equation}
%
This is useful in numerical maximisation.

>>>>

Given a sequence of plasticity events at times $t_1,t_2,t_3,\ldots$, the probability distribution across the internal states at time $t_n$ is
%
\begin{equation}\label{eq:plastseq}
  \pr(t_n) = \pr(t_0)  \M\pot \M\dep \M\dep \M\pot \M\dep \M\pot \M\pot \M\dep \ldots \M\pot .
\end{equation}
%
If the probability of a given plasticity event being potentiating/depressing is $f\potdep$, the probability of the sequence above (conditioned on the number of events) is given by $f\pot f\dep f\dep f\pot f\dep f\pot f\pot f\dep \ldots f\pot$.

@@

This results in
%
\begin{equation}\label{eq:overlapwt}
\begin{aligned}
  \av{\syn^1_\text{ideal}\syn^1(t)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \e^{rt\frg} \w, \\
  \av{\syn^1_\text{ideal}\syn^1(\infty)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \onev\eq \w \\
         &= \eq (f\pot\onev-f\dep\onev)\, \eq \w \\
         &=  (f\pot-f\dep)\, \eq \w \\
         &=  (f\pot-f\dep)\, \eq \e^{rt\frg} \w .
\end{aligned}
\end{equation}
%
Combining these allows us to write the numerator of \eqref{eq:SNRdef} as
%
\begin{equation}\label{eq:signal}
\begin{aligned}
  \av{\syn_\text{ideal}\cdt\syn(t)} - \av{\syn_\text{ideal}\cdt\syn(\infty)}
    &= N \eq \enc \, \e^{(r\frg+\hom)t} \w , \\
    \text{where} \qquad
  \enc &= f\pot(\M\pot-\I)-f\dep(\M\dep-\I)
\end{aligned}
\end{equation}
%
Combining with \eqref{eq:noise} gives
%
\begin{equation}\label{eq:SNRcurveExact}
  \SNR(t) = \frac{\sqrt{N} \eq \enc\,\e^{rt\frg}\w}
                 {\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
\end{equation}
%
%We can absorb $\hom$ into $\M\potdep$ by defining
%%
%\begin{equation}\label{eq:absorbhom}
%\begin{aligned}
%  \Mh\potdep &= \M\potdep + \frac{\hom}{2rf\potdep},
%  &\qquad
%  \frgh &= f\pot\Mh\pot + f\dep\Mh\dep - \I,
%  \\&&
%  \ench &= f\pot(\Mh\pot-\I)-f\dep(\Mh\dep-\I).
%\end{aligned}
%\end{equation}
%%
%Then the SNR curve is
%%
%\begin{equation}\label{eq:SNRcurvehom}
%  \SNR(t) = \frac{\sqrt{N}  \eq \enc \,\e^{rt\frg}\w}
%                 {\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
%\end{equation}
%%
The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with $f\pot=f\dep$ or $\eq_+=\eq_-$.
We can perform the maximisation in two steps.
First maximise the numerator at fixed $\eq\w$.
Then maximise the ratio \wrt $\eq\w$.
This will allow us to ignore the denominator until the very end.

We can then treat the off-diagonal elements of $\M\potdep$ as the independent degrees of freedom, subject to these constraints, when maximising the SNR.
The diagonal elements are determined by setting the row sums to 1.
We then have
%
\begin{equation}\label{eq:derivpd}
  \pdiff{\M\pot_{ij}}{\M\pot_{mn}} = \pdiff{\M\dep_{ij}}{\M\dep_{mn}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
  \qquad
  \pdiff{\M\pot_{ij}}{\M\dep_{mn}} = \pdiff{\M\dep_{ij}}{\M\pot_{mn}} = 0.
\end{equation}
%
It will also be convenient to define the differential operators
%
\begin{equation}\label{eq:pertfe}
  \pdiff{}{\frg_{ij}} = \frac{1}{2f\pot} \pdiff{}{\M\pot_{ij}} + \frac{1}{2f\dep} \pdiff{}{\M\dep_{ij}},
  \qquad
  \pdiff{}{\enc_{ij}} = \frac{1}{2f\pot} \pdiff{}{\M\pot_{ij}} - \frac{1}{2f\dep} \pdiff{}{\M\dep_{ij}},
\end{equation}
%
these satisfy
%
\begin{equation}\label{eq:derivfe}
  \pdiff{\frg_{ij}}{\frg_{mn}} = \pdiff{\enc_{ij}}{\enc_{mn}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
  \qquad
  \pdiff{\frg_{ij}}{\enc_{mn}} = \pdiff{\enc_{ij}}{\frg_{mn}} = 0.
\end{equation}
%

@@

Now consider the Laplace transform of the SNR curve \eqref{eq:SNRcurveExact} (ignoring the denominator)
%
\begin{equation}\label{eq:lsnr}
\begin{aligned}
  A(s) &= \intd[_0^{\infty}]{t} \e^{-st}\SNR(t) \\
   &= \intd[_0^{\infty}]{t} \frac{ \sqrt{N} \eq \enc \e^{(r\frg-s)t} \w }{ \sqrt{1-(f\pot-f\dep)^2(\eq\w)^2} } \\
   &= \frac{\sqrt{N} \eq \enc \prn{s - r\frg}\inv \w }{\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
\end{aligned}
\end{equation}
%
This expression is ill-behaved at $s=0$.
Thanks to \eqref{eq:alfund}, we can solve this by the replacement $\mathbf{G}(s) \to \fund(s)$, as $\enc\onev=0$.
%
\begin{equation}\label{eq:larea}
\begin{aligned}
  \hat{A}(s) = A(s) \sqrt{\frac{{{1-(f\pot-f\dep)^2(\eq\w)^2}}}{N}}  &=  \eq \enc \fund(s) \w \\
    &= \sum_{ijk} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \eq_k \w_k \\
    &= \sum_{ij}  \eq_i \enc_{ij} (\etw_{i}(s) - \etw_{j}(s)).
\end{aligned}
\end{equation}
%
Note that $A(0)=A$, the area, and $\lim_{s\to\infty} \brc{sA(s)} = \SNR(0)$, the initial SNR.

In analogy with the case $s=0$, we define
%
\begin{equation}\label{eq:lareacoeffs}
  \begin{aligned}
    c_k(s) &= \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)}, \\
    a_j(s) &= \sum_{jk} \eq_j \enc_{jk} (\etw_{j}(s) - \etw_{k}(s)), \\
    \thb_i(s) &= \sum_j \fptb_{ij}(0)\ a_j(s), \\
  \end{aligned}
    \implies
    \hat{A}(s) = \sum_k c_k(s) \eq_k \w_k
      = \sum_i a_i(s).
\end{equation}
%

@@

Using the same approach as the area bound in \cite{Lahiri2013synapse} we find that the derivatives are
%
\begin{equation}\label{eq:lAdiff}
  \pdiff{\hat{A}(s)}{\M\potdep_{mn}}
     = f\potdep \eq_m \prn{ r \brk{\thb_m(s) - \thb_n(s)}
     +\prn{rc_m(s) \pm 1} \brk{\etw_m(s) - \etw_n(s)} },
\end{equation}
%
with the upper/lower sign in $\pm$ corresponding to derivatives \wrt $\M\potdep$.
The term with $\Theta(s)$ comes from differentiating $\eq$, the term with $c(s)$ comes from differentiating $\etw(s)$ and the term with $\pm1$ comes from differentiating $\enc$ in \eqref{eq:larea}.

To write down the Hessian, we will first define some notation.
We define the matrix $\mathbf{X}(s) = \fund(0) \enc \fund(s)$.
Given a four-index array $H_{ijkl}$, we define $H\trans_{ijkl} = H_{klij}$.
Then we denote
%
\begin{equation}\label{eq:hessparts}
\begin{aligned}
  R_{ijkl} &=
    &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\thb_k(s) - \thb_l(s)} \\
    &&&+ \eq_i \brc{\mathbf{X}_{ik}(s) - \mathbf{X}_{jk}(s)} \brc{\etw_k(s) - \etw_l(s)} \\
    &&&+ \eq_i c_i(s) \brc{\fund_{ik}(s) - \fund_{jk}(s)} \brc{\etw_k(s) - \etw_l(s)},  \\
  S_{ijkl} &=
    &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\etw_k(s) - \etw_l(s)} \\
    &&&+ \eq_k \brc{\fund_{ki}(s) - \fund_{li}(s)} \brc{\etw_i(s) - \etw_j(s)}.
\end{aligned}
\end{equation}
%
Then we can write the Hessian as
%
\begin{equation}\label{eq:hessian}
\begin{aligned}
  \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\pot_{kl}} &=
     (f\pot)^2   [R + R\trans + S + S\trans]_{ijkl}, \\
  \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\dep_{kl}} &=
     f\pot f\dep [R + R\trans - S + S\trans]_{ijkl}, \\
  \pdiff{^2\hat{A}(s)}{\M\dep_{ij} \partial\M\pot_{kl}} &=
     f\pot f\dep [R + R\trans + S - S\trans]_{ijkl}, \\
  \pdiff{^2\hat{A}(s)}{\M\dep_{ij} \partial\M\dep_{kl}} &=
     (f\dep)^2   [R + R\trans - S - S\trans]_{ijkl}, \\
\end{aligned}
\end{equation}
%
This is useful in numerical maximisation.
